\documentclass[final,a4paper,twosided]{thenst}

\usepackage{url}
\usepackage{listings}
\usepackage{pslatex}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[cyr]{aeguill}
\usepackage[dvips]{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{pstricks}
\usepackage{verbatim}
\usepackage{algorithm2e}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[english,francais]{babel}
%\usepackage[scaled]{helvet}
%\renewcommand{\familydefault}{\sfdefault}

\let\urlorig\url
\renewcommand{\url}[1]{
  \begin{otherlanguage}{english}\urlorig{#1}\end{otherlanguage}%
}

\lstdefinelanguage{aadl} {
  morekeywords={in,out,package,end,bus,data,thread,port,group,process,processor,
    system,memory,device,subprogram,public,private,event,property,set,applies,to,
    units,type,implementation,parameter, reference},
  morekeywords={properties,features,annex,modes,connections,flows,
    subcomponents,calls,binding},
  morekeywords={aadlinteger,aadlboolean,aadlstring,aadlfloat},
  morecomment=[l]{--},
}

\lstdefinelanguage{lustre} {
  morekeywords={if, then, else, pre, node, returns, let, tel, int,
    bool, real, var, when},
  morekeywords={true, false}
}

\lstdefinelanguage{lotos} {
  morekeywords={specification, noexit, exit, behavior, process,
    endproc, endspec, stop, where, eq, NAT, ne, and, or, neq},
}

\lstset {
  language=ada,
  basicstyle=\footnotesize\sffamily,
  aboveskip=\bigskipamount,
  belowskip=\smallskipamount,
  abovecaptionskip=\smallskipamount,
  belowcaptionskip=\medskipamount,
  xleftmargin=\parindent,
  captionpos=b,
  tabsize=2,
  frame=single
}

\newcommand{\Concept}[1]{#1\xspace}
\newcommand{\Project}[1]{#1\xspace}

\renewcommand{\baselinestretch}{1}
\newcommand{\listingwidth}{0.95\linewidth}
\newcommand{\todo}[1]{{\bf \emph{TODO~#1}}}
\newcommand{\addref}[1] {\todo{Ref on : #1}}
\newcommand{\aadl}{AADL\xspace}
\newcommand{\ada}{Ada\xspace}
\newcommand{\newada}{\Concept{Ada~2005}}
\newcommand{\rcm}{RCM\xspace}
\newcommand{\hrtuml}{HRT-UML\xspace}
\newcommand{\rsprof}{Ravenscar Profile\xspace}
\newcommand{\hi}{HI\xspace}
\newcommand{\osate}{\Project{OSATE}}
\newcommand{\assert}{ASSERT\xspace}
\newcommand{\norb}{nORB}
\newcommand{\microorb}{microORB}
\newcommand{\vm}{VM\xspace}
\newcommand{\polyorbhi}{PolyORB-HI\xspace}
\newcommand{\polyorb}{PolyORB\xspace}
\newcommand{\ocarina}{Ocarina\xspace}
\newcommand{\gnat}{GNAT\xspace}
\newcommand{\glade}{GLADE\xspace}
\newcommand{\corba}{CORBA\xspace}
\newcommand{\stood}{Stood\xspace}
\newcommand{\hood}{HOOD\xspace}
\newcommand{\kw}[1]{\texttt{\textbf{#1}}}
%\newcommand{\dataport}{\pspolygon[linewidth=1pt,fillcolor=black,fillstyle=solid]
%(0,0)(0,0.2)(0.2,0.1)\xspace\xspace\xspace}
%\newcommand{\eventport}{\psline[linewidth=1pt](0,0)(0.2,0.1)(0,
%0.2)\xspace\xspace\xspace}
%\newcommand{\eventdataport}{\pspolygon[linewidth=1pt,fillcolor=black,
%fillstyle=solid](0.02,0.07)(0.02,0.13)(0.08,0.1)\psline[linewidth=0.5pt](0,
%0)(0.2,0.1)(0 ,0.2)\xspace\xspace\xspace}
\newcommand{\dataport}{}
\newcommand{\eventport}{}
\newcommand{\eventdataport}{}
\newcommand{\ts}{$\tau_s$\xspace}
\newcommand{\tl}{$\tau_l$\xspace}
\newcommand{\plong}{$T_{long}$\xspace}
\newcommand{\pshort}{$T_{short}$\xspace}
\newcommand{\simu}{Simulink$^\circledR$}
\newcommand{\susp}{{\tiny$\ldots$}\xspace}
\newcommand{\clearemptydoublepage}{\newpage{\pagestyle{empty}
    \cleardoublepage}}
\newcommand{\MM}[1]{\ensuremath{#1}}
\newcommand{\REGLE}[3]{
%  \box{
    \MM{
      \dfrac{\begin{array}{l} #1 \end{array}}
	    {\begin{array}{l} #2 \end{array}}
    }
%  }
  %\vspace{1mm}
    \begin{left}
      \hbox{#3}
    \end{left}
}               
\newcommand{\regle}[2]{
  \MM{
    \frac{\begin{array}{c} #1 \end{array}}
         {\begin{array}{l} #2 \end{array}}
  }
}
\newcommand{\formatreg}[1]{
  \begin{center}
%  \begin{tabular}{|c}
%    \hline
    \vspace{0.5mm}
    \fbox{#1}\\
%    \hline
%  \end{tabular}
  \end{center}
}
\newcommand{\Fait}[1]{\MM{\ \ \xrightarrow{#1}}}
\newcommand{\fait}[1]{\MM{\xrightarrow{\ \ #1}}}
\newcommand{\Faitv}[1]{\MM{\ \xrightarrow{#1}\ }}
\newcommand{\RL}[1]{\,\frac{}{\,^{{#1}}\,}\,}
\newcommand{\RS}[1]{\,\frac{}{^{{#1}}}\,}
\newcommand{\sqplus}{\ {\scriptstyle {\widehat{+}}}\ }
\newcommand{\smplus}{\ {\scriptstyle {\dot{+}}}\ }
\newcommand{\intrupt}{\,{\scriptstyle\leadsto}\,}
\newcommand{\rget}{\frac{}{\text{\ttfamily\ Get\ }}}
\newcommand{\rset}{\frac{}{\text{\ttfamily\ Set\ }}}
\newcommand{\rgvt}{\MM{ \frac{}{\text{\ttfamily\  Await\_Event\ }} }}
\newcommand{\rsvt}{\frac{}{\text{\ttfamily\ Send\_Event\ }}}
%\newcommand{\mget}{\frac{}{{\scriptsize{\text{\ttfamily\ Get\ }}}}
%\newcommand{\mset}{\frac{}{\text{\ttfamily\ Set\ }}}
\newcommand{\mgvt}{\frac{}{\text{\ttfamily\ \ Await\_Event\ }}}
\newcommand{\msvt}{\frac{}{{\text{\ttfamily\ \tiny Send\_Event\ }}}}
\newcommand{\rx}{\frac{}{\large \ x \ }}
\newcommand{\smcirc}{{\scriptscriptstyle \circ}}
\newcommand{\smodot}{{\scriptscriptstyle \odot}}
\newcommand{\smoplus}{\,{\scriptstyle \oplus}\,}
\newcommand{\SFT}[1]{\text{\sffamily #1}}
\newcommand{\FNT}[1]{\text{\footnotesize #1}}
\newcommand{\SC}[1]{$\text{\scshape{#1}}$}


\restylealgo{algoruled}

\author{Irfan Hamid}
\supervisor{Elie Najm}
\president{Nicolas Halbwachs}
\committee{
  Tullio Vardanega & Reviewer\\
  Jean-Fran\c cois Pradat-Peyre & Reviewer\\
  Peter H. Feiler & Examiner\\
  Juan Antonio de le Puente & Examiner
}
\thesisdate{30$^{th}$ of May, 2008}
\title{Automatic Code Generation and Verification of Hard Real-time
  Systems}

\frtitle{G\'en\'eration automatique du code et la v\'erification des
  syst\`emes temps-r\'eels}

\enabstract{The increasing complexity of real-time software and a
  corresponding decrease in times-to-market has necessitated new
  software development methods for such systems, which expedite
  various aspects of their engineering. Additionally, these systems
  are often safety-critical, where a fault in the software could
  result in loss of human life, which leads to strong correctness and
  safety requirements being imposed on such software and associated
  development processes. In this thesis, a new approach is explored
  for the automatic generation of source code for real-time systems
  from an architecture description language (ADL). The ADL chosen is
  the Architecture Analysis \& Design Language, which is specific to
  the real-time domain. The target runtime chosen is the Ravenscar
  Profile of the Ada language for high-integrity systems, which is
  robust and suitable for high-integrity systems and offline
  schedulability analysis. The system description in the AADL is
  converted to an ``execution framework'' which provides the runtime
  entities such as threads and communication buffers corresponding to
  the application architecture. This execution framework is then
  fleshed out with the functional behavior of threads---in the form of
  callbacks---to give a complete system. The code generation also
  includes a verified solution for the control system problem of
  deterministic inter-task communication. A formal semantics of the
  target language is provided, thus allowing for the verification of
  target code. A code generation tool has been developed that
  implements the approach. Finally, a major avionics case study is
  re-engineered using the developed tooling to validate the approach
  for the considered domain.}

\frabstract{La complexité croissante des logiciels temps-réels ainsi
  que la réduction du temps de leur mise en marché necessitent le
  développement de nouvelles méthodes afin d'accélérer divers aspects
  d'ingénierie de ces systèmes. De plus, ces derniers sont souvent
  critiques, puisqu'une défaillance dans le logiciel peut engendrer
  des dégâts humains, ce qui implique de fortes exigences en matière
  de sécurité. Nous avons exploré dans le cadre de cette thèse, une
  nouvelle approche pour la génération automatique de code source pour
  les systèmes temps-réels à partir d'un langage de déscription
  d'architecture (ADL pour Architecture Description Language). L'ADL
  choisi est le Architecture Description and Design Language (AADL),
  spécifique au domaine temps-réel. Nous avons choisi d'utiliser le
  profil Ravenscar du langage Ada; ce profil est un sous-ensemble
  d'Ada dédié au systèmes temps-réels et critique, il est robuste et
  se donne facilement à l'analyse d'ordonnonçabilité hors-ligne. La
  déscription du système en AADL est convertie en un cadre d'execution
  qui fournit les entitiés d'execution tels que les threads et les
  tampons de communication correspondant à l'architecture de
  l'application. Ensuite, ce cadre d'execution est complété par des
  comportements fonctionnels des threads (sous formes de
  callbacks). La génération de code inclut également une solution au
  problème de communication déterministe entre les threads, un
  problematique qui se présente dans le domaine d'automatisme; aussi
  fournit est une méthode de généré de la vérification formelle pour
  ces tampons de communication déterministe. Une sémantique formelle
  du langage sible est fournie, permettant de vérifier le code
  cible. Nous avons développé un outil de génération implementant
  notre approche. Enfin, nous avons validé notre approche en
  l'applicant à une étude de cas avionique bien détaillée.}

\begin{document}
\begin{otherlanguage}{english}
\maketitle

\begin{dedication}
For my parents
\end{dedication}
\end{otherlanguage}

\begin{acknowledgement}
No man is an island; this is even truer for an endeavour such as
research, where we all stand on the shoulders of giants. A plethora of
people have helped and influenced me during my doctorate at the
Computer Science and Networks Department of the ENST. First among
these is, without a doubt, my advisor Elie Najm, whom I wish to thank
for so much, including his guidance, patience, encouragement and
friendship, for without him I would not have been able to complete my
doctorate.

My colleagues at the ENST, an exemplary set of people, who became dear
friends over the course of my stay here. Raja {\sc Chiky}, my office
colleague of 3 years now; Khaled {\sc Barbaria}, my other office
colleague and a good friend and sometimes mentor of Ada. Bechir {\sc
  Zalila}, a kindred soul who worked on code generation as well, and
one of the most selfless people---and one of the best programmers---I
know, always ready to share ideas as well as a coffee. Hoa {\sc Ha
  Duong}, the walking encyclopedia to whom I shall always be grateful
for---among other things---having introduced me to \emph{Doctor Who}
and \emph{The Big Bang Theory}. Laurent {\sc Pautet}, who although not
my advisor, always found the time to help me out whenever I needed,
and to share a coffee and a smoke. Thomas {\sc Vergnaud}, the original
``Mr. AADL'' of our little modeling and code generation group. Ahmad
{\sc Fadlallah} for his help and company. Olivier {\sc Gilles}, with
whom I've had endless discussions on (geo)politics, and Isabelle {\sc
  Perseil}, with whom I've had endless discussions on modeling. Jérôme
{\sc Hugues} for his incisive and always pertinent advice and
admonishments. Julien {\sc Delange}, who will soon be the Linus {\sc
  Torvalds} of the real-time systems community and Gilles {\sc
  Lasnier}, the bon-vivant who is always cool as a cucumber and more
than ready to lend a hand, I wish I had known you both
earlier. Muhammad {\sc Tahir}, my compatriot in the
department. Isabelle {\sc Demeure}, the head of the department and one
of the gentlest people with the readiest smile I've had the pleasure
of knowing. The very helpful, generous and \emph{always} boisterous
and rambunctious department administrator Hayette {\sc Soussou}, as
well as Celine {\sc Bizart}.

My friends from Pakistan who came with me for studies to
France. Specially the ones from my old engineering school: Asad,
Usman, OK, Taj, Bilal, Kiani, Saadi, Cheema, Khattak. They say that
``friends help you move, good friends help you move bodies'', you guys
are \emph{great} friends.

My friends from my Master of Science days in Paris. Ruben and
Maria-Eugenia, a wonderful couple and a true testament to the merits
of getting married. Kapil and Eftychia, a formidable man and his
equally fiery woman. Rana and Wenping, always caring and ready to
help. Aida, the twinkly-eyed Spanish pixie who is lost to us somewhere
in Switzerland. My fellow geek and neighbor Mevegni. The ``Prince of
Persia'' Hossein. Paco, the ``real'' researcher over at JHU. So many
others: Sanaz, Victor, Claire, Marios, Alberto, Laura \ldots

I had the good luck of staying at the Maison d'Italie during the
entirety of my doctoral studies where I got to know some special
people. Avais {\sc Daulat}, the lively neighbor whose door was always
open and whom I consider myself lucky to have known, not to mention
his better half Eleonora. Olivia and Sebastien, always welcoming and
ready to have fun. Giulia and Marco, who came from Italy to attend my
dissertation. Carolina, Krystelle, Sandra, Nadja, Celine, Naomi,
Darius, Elena, Abdel, Teresa, Francesca, Maurice, Shin, Fumi, Kurtis,
Roberta \ldots

Finally, and most importantly, my parents Insha and Asiya, who have
always nurtured my curiosity. It is to them that I owe all for they
have always encouraged me to do whatever I wanted, so thank
you. Bilal, the best younger brother a man could hope for, and then
some.
\end{acknowledgement}

\makeabstracts

\begin{otherlanguage}{francais}
% Use only asterisked versions of section/subsection commands %
\input{french_summary}
\cleardoublepage%Necessary to preserve hdr/ftr on last page of summary
\end{otherlanguage}

\begin{otherlanguage}{english}
\setcounter{tocdepth}{2}
\tableofcontents
\listoftables
\listoffigures

%\input{glossary}

%\input{intro.tex}
%\input{biblio.tex}
%\input{aadlrs.tex}
%\input{code_gen.tex}
%\input{adv_gen.tex}
%\input{formal_sem.tex}
%\input{conclusions.tex}

\bibliographystyle{plain}
\bibliography{hamid}

\appendix
\input{mcc.tex}
\input{ada.tex}
\end{otherlanguage}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% mode: flyspell
%%% TeX-master: t
%%% End:
